Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-21 manuscript/manuscript.tex
Prepare for submission of the manuscript
Convert to single column layout and fix the size of some of the figures.
Fixed grid-search figure caption
Removed citation in conclusions

\documentclass[extra]{gji}
%~ \documentclass[extra, referee]{gji}

\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{todonotes} % to make annotations on margins
%~ \usepackage[disable]{todonotes} % to disable annotations on margins

@@

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/tesseroid-uieda.pdf}
\caption{
    A tesseroid (spherical prism) in a geocentric spherical coordinate system, with a
    computation point $P$ and its local north oriented Cartesian coordinate system.
    After \citet{Uieda2015}.
}
\label{fig:tesseroid}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]
    {figures/grid-search.pdf}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison
    between the analytical solution and the numerical approximation of
    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
    These comparisons were carried out on the ``Big Grid'' configuration (Table
    \ref{tab:grids}), with a spherical shell of
    35km of thickness, and density function with $b$ of 1km.
    The points inside the dashed line are the ones that present an
    error lower than 0.1\%.
    }
\label{fig:grid-search}
\end{figure}

@@

Because there is no direct relation between the values assigned to $D$ and $\delta$ and
the error of the computation, we empirically determined the optimum values for these
parameters.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimum values were a linear and an
exponential function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
We analyzed the error for exponential functions ranging from smooth to sharp variations
in depth to test the accuracy of the density-based discretization.
Through this analysis, we have obtained optimal values for the distance-size ratio of
1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
respectively.
These values are in agreement with previous results for a homogeneous density tesseroid
\citep{Uieda2016}.
The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
0.1\% precision.
A wide range of exponential functions were used to infer these results and they are
conservative estimates, chosen to prioritize accuracy over speed of computations.
Therefore, they can be safely generalized to other common depth-density relations, like
quadratic functions, without loss of accuracy.

@@

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
\caption{
    Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
    The computation point $Q$ is located on the $z$ axis at a distance $r$ from
    the origin of the coordinates system.
    For our purposes we will assume that $Q$ is outside of the shell,
    i.e. $r > R_2$.
}
\label{fig:spherical-shell}
\end{figure}

@@

\iftwocol{
\begin{equation}
    \begin{split}
        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
        Ab
        \Big[
        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - R}{b}} - \\
        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - R}{b}}
        \Big] + \\
        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
    \end{split}
\label{eq:shell-pot-exp}
\end{equation}
}{
\begin{equation}
    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
    Ab \, e^\frac{R}{b}
    \Big[
    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} -
    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
    \Big] +
    \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
\label{eq:shell-pot-exp}
\end{equation}
}

>>>>

%\documentclass[extra]{gji}
\documentclass[extra, referee]{gji}

\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}

@@

\begin{figure}
\centering
\includegraphics[width=0.6\linewidth]{figures/tesseroid-uieda.pdf}
\caption{
    A tesseroid (spherical prism) in a geocentric spherical coordinate system, with a
    computation point $P$ and its local north oriented Cartesian coordinate system.
    After \citet{Uieda2015}.
}
\label{fig:tesseroid}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]
    {figures/grid-search.pdf}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison
    between the analytical solution and the numerical approximation of
    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
    These comparisons were carried out on the ``Big Grid'' (Table
    \ref{tab:grids}), with a spherical shell of
    35km of thickness, and density function with $b$ of 1km.
    The points inside the dashed line are the ones that present an
    error lower than 0.1\%.
    }
\label{fig:grid-search}
\end{figure}

@@

Because there is no direct relation between the values assigned to $D$ and $\delta$ and
the error of the computation, we empirically determined the optimum values for these
parameters.
These values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimum values were a linear and an
exponential function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
We analyzed the error for exponential functions ranging from smooth to sharp variations
in depth to test the accuracy of the density-based discretization.
Through this analysis, we have obtained optimal values for the distance-size ratio of
1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
respectively.
These values are in agreement with previous results for a homogeneous density tesseroid.
The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
0.1\% precision.
A wide range of exponential functions were used to infer these results and they are
conservative estimates, chosen to prioritize accuracy over speed of computations.
Therefore, they can be safely generalized to other common depth-density relations, like
quadratic functions, without loss of accuracy.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{figures/spherical-shell.pdf}
\caption{
    Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
    The computation point $Q$ is located on the $z$ axis at a distance $r$ from
    the origin of the coordinates system.
    For our purposes we will assume that $Q$ is outside of the shell,
    i.e. $r > R_2$.
}
\label{fig:spherical-shell}
\end{figure}

@@

\begin{equation}
    \begin{split}
        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
        Ab
        \Big[
        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - R}{b}} - \\
        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - R}{b}}
        \Big] + \\
        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
    \end{split}
\label{eq:shell-pot-exp}
\end{equation}
